---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

anonymize <- function(df) {
  drop <- c(
    "PRJ_LDR.USERNAME", "PRJ_LDR.FIRST_NAME",
    "PRJ_LDR.LAST_NAME", "PRJ_LDR"
  )
  drop <- c(drop, tolower(drop))
  df <- df[, !(names(df) %in% drop)]
}
```
# glfishr

*glfishr* contains a series of R functions that are intended to make
it easy to get fisheries assessment and creel survey data from the
fn_portal and creel_portal api's and into R for subsequent analysis
and reporting.  Functions are named semantically to reflect the FN-II
table they fetch data from.  There are functions that are specific to
assessment programs such as `get_FN011()` for project meta data,
`get_FN121()` for net set/sample data and `get_FN125()` for biological
sample data.  There are analogous functions to fetch data for creels:
`get_SC011()` for creel survey meta data, `get_SC121()` for creel
survey interview records and `get_SC125()` for biological data
collected from fish sampled in creels.  Most of the functions take an
optional filter_list parameter that can be used to finely control
which records are returned.  Care has been taken to ensure that the
available filters are consistent with FN-II field names whenever
possible, and in many cases, filters can be re-used across different
tables (e.g. - if a filter is passed to the `get_FN121()` function to
find a subset of net sets, that same filter can be applied to the
`get_FN125()` function to get the all of the biological samples
collected in those net sets).

All of the filters are specified using the following convention:
`<field_name>__<expression>` - the field name, two underscores, and
the expression that is to be applied to the field to select the subset
of records.  In most cases, the field name is a lowercase fishnet-II
field (prj_cd, tlen, gon, ect.) and will be documented in the Data Dictionary.
The available expressions are dependent on the type of data in the field. In
most cases, strings (such as prj_nm, prj_cd) can be filtered with 'like',
'not_like', and 'endswith'.  Fields with a well defined number of choices
(prj_cd, spc, gon, sex, tagstat)  can be selected by passing a
comma separated list of choices to include or exclude (`gon=10,20` or
`gon__not=10,20`), as well as null or not_null.  Numeric fields (sidep,
flen, rwt etc) typically have the following filter expressions available:

+ equal: `year=2010`
+ greater than or equal to: `tlen__gte=350`
+ greater than: `tlen__gt=350`
+ less than or equal to: `flen__lte=350`
+ less than: `flen__lt=350`
+ null: `clipc__null`
+ not null: `rwt__not_null`

Many of these filters are illustrated in the following examples, more
detailed infromation can be found using the `show_filters()` function
that takes a table name, and optionally, a partial filter name to
match against.  `show_filters()` will print out all of the filters
available for that table and, if appropriate, provide additional
information on expected format (eg. "format: yyyy-mm-dd").  If no
endpoint or table name is provided, `show_filters()` will display the
names of the filterable endpoints that are currently available.

Most of the functions accept a list of filters that can be used to
select the data that is returned from the Great Lakes Stocking
Database.  The list forms a key-value pair that specified the filter
to be applied and the value of the filter.


```{r show_filters }
show_filters()

show_filters("fn011")

show_filters("fn125", filter_like = "tlen")
```



## Load glfishr

All of the functions in glfishr have been bundled up into an R-package
that can be installed and then loaded as needed:

```{r load_library}
library(glfishr)
```


## FN011 - Projects

Project meta data can be accessed using the `get_fn011()` function.
FN011 records contain the hi-level meta data about an
OMNR netting project.  The FN011 records contain information like
project code, project name, project leader, start and end date,
protocol, and the lake where the project was conducted.  This
function takes an optional filter list which can be used to select
records based on several attributes of the project such as
project code, or part of the project code, lake, first year (**year__gte**), last
year (**year__lte**), protocol, etc.


```{r fn011}
fn011 <- get_FN011(list(lake = "ON", year__gte = 2012, year__lte = 2018))
fn011 <- anonymize(fn011)
nrow(fn011)
head(fn011)

fn011 <- get_FN011(list(lake = "ER", protocol = "TWL"))
fn011 <- anonymize(fn011)
nrow(fn011)
head(fn011)


filters <- list(lake = "SU", prj_cd = c("LSA_IA15_CIN", "LSA_IA17_CIN"))
fn011 <- get_FN011(filters)
fn011 <- anonymize(fn011)
nrow(fn011)
head(fn011)


fn011 <- get_FN011(list(lake = "HU", prj_cd__like = "_006"))
fn011 <- anonymize(fn011)
nrow(fn011)
head(fn011)
```


## FN121 - Net Sets

Net sets can be retrieved using the `get_FN121()` function.  The FN121
records contain information like set and lift date and time, effort
duration, gear, site depth and location.  This function takes an
optional filter list which can be used to return records based on
several attributes of the net set including set and lift date and
time, effort duration, gear, site depth and location as well as
attributes of the projects they are associated with such as project code
or part of the project code, lake, first year, last year, protocol,
etc.


```{r fn121}
fn121 <- get_FN121(list(lake = "ON", year = 2012))
nrow(fn121)
head(fn121)


fn121 <- get_FN121(list(lake = "ER", protocol = "TWL", year__gte = 2010, sidep__lte = 20))
nrow(fn121)
head(fn121)


filters <- list(
  lake = "SU",
  prj_cd = c("LSA_IA15_CIN", "LSA_IA17_CIN")
)

fn121 <- get_FN121(filters)

nrow(fn121)
head(fn121)


fn121 <- get_FN121(list(lake = "HU", prj_cd__endswith = "_003"))
nrow(fn121)
head(fn121)
```


## FN122 - Sample Efforts


Sample Efforts can be retrieved using the `get_fn122()` function. FN122
records contain information about efforts within a sample.  For most
gill netting project an effort corresponds to a single panel of a
particular mesh size within a net set (gang). For trap netting and
trawling projects, there is usually just a single effort. The FN122
table contains information about that particular effort such as gear
depth, gear temperature at set and lift, and effort distance.  This
function takes an optional filter list which can be used to return
records based on several attributes of the effort including effort
distance and depth but also attributes of the projects or nets set
they are associated with such as project code, lake, first year, last
year, protocol, gear etc.


```{r fn122}
fn122 <- get_FN122(list(lake = "ON", year = 2012, gear = "GL", sidep__lte = 15))
nrow(fn122)
head(fn122)


filters <- list(
  lake = "ER",
  protocol = "TWL",
  year__gte = 2010,
  sidep__lte = 20
)
fn122 <- get_FN122(filters)
nrow(fn122)
head(fn122)


filters <- list(
  lake = "SU",
  prj_cd = c("LSA_IA15_CIN", "LSA_IA17_CIN"), eff = "051"
)
fn122 <- get_FN122(filters)
nrow(fn122)
head(fn122)



filters <- list(lake = "HU", prj_cd__like = "_007", eff = c("127", "140"))
fn122 <- get_FN122(filters)
nrow(fn122)
head(fn122)
```



## FN123 - Catch Counts


Catch counts by effort, species, and group are available using the
`get_FN123()` function.  FN123 records contain information about catch
counts by species for each effort in a sample.  For most gill netting
projects this corresponds to catches within a single panel of a
particular mesh size within a net set (gang). Group (GRP) is
occasionally included to further sub-divide the catch into user defined
groups that are usually specific to the project, but will always be
included and will be '00' by default. This function takes an optional
filter list which can be used to return records based on several
attributes of the catch including species or group code, but also
attributes of the effort, the sample or the project(s) that the
catches were made in.


```{r fn123}
fn123 <- get_FN123(list(lake = "ON", year = 2012, spc = "334", gear = "GL"))
nrow(fn123)
head(fn123)

filters <- list(
  lake = "ER",
  protocol = "TWL",
  year = 2010,
  spc = c("331", "334"),
  sidep__lte = 20
)
fn123 <- get_FN123(filters)
nrow(fn123)
head(fn123)


filters <- list(
  lake = "SU",
  prj_cd = c("LSA_IA15_CIN", "LSA_IA17_CIN"),
  eff = "051",
  spc = "091"
)
fn123 <- get_FN123(filters)
nrow(fn123)
head(fn123)


filters <- list(lake = "HU", spc = "076", grp = "55")
fn123 <- get_FN123(filters)
nrow(fn123)
head(fn123)
```



## FN124 - Length Tallies

An api endpoint and associated function for FN124 records has not been created yet, but will be coming soon.



## FN125 - Biological Data

Biological data is maintained in the FN125 table and can be accessed
using the `get_FN125` function. FN125 records contain the biological
data collected from individual fish sampled in assessment projects
such as length, weight, sex, and maturity. For convenience this end point
also returns data from child tables such as the 'preferred' age, and
lamprey wounds.  This function takes an optional filter list which can
be used to return records based on several different biological
attributes (such as size, sex, or maturity), but also of the species,
or group code, or attributes of the effort, the sample, or the
project(s) that the samples were collected in.

```{r fn125}
fn125 <- get_FN125(list(lake = "ON", year = 2012, spc = "334", gear = "GL"))
nrow(fn125)
head(fn125)

filters <- list(
  lake = "ER",
  year = "2019",
  protocol = "TWL",
  spc_in = c("331", "334"),
  sidep__lte = 10
)
fn125 <- get_FN125(filters)
nrow(fn125)
head(fn125)

filters <- list(
  lake = "SU",
  prj_cd = c("LSA_IA15_CIN", "LSA_IA17_CIN"),
  eff = "051",
  spc = "091"
)
fn125 <- get_FN125(filters)
nrow(fn125)
head(fn125)



filters <- list(lake = "HU", spc = "076", grp = "55")
fn125 <- get_FN125(filters)
nrow(fn125)
head(fn125)
```


## FN125Tags - Tags Recovered or Applied

FN125Tags records contain information about the individual tags
applied to or recovered from on a sampled fish and can be fetched from
the api using `get_FN125Tags()` function.  Historically, tag data was
stored in three related fields - TAGDOC, TAGSTAT and TAGID.  This
convention is fine as long a single biological sample only has a one
tag. In recent years, it has been come increasingly common for fish to
have multiple tags, or tag types associated with individual sampling
events. FN125Tag accommodates those events.  This function takes an
optional filter list which can be used to return records based on
several different attributes of the tag (tag type, colour, placement,
agency, tag stat, and tag number) as well as, attributes of the
sampled fish such as the species, or group code, or attributes of the
effort, the sample, or the project(s) that the samples were collected
in.


```{r fn125Tags}
fn125_tags <- get_FN125Tags(list(
  lake = "ON",
  year = 2019,
  spc = "081",
  gear = "GL"
))
nrow(fn125_tags)
head(fn125_tags)


fn125_tags <- get_FN125Tags(list(lake = "SU"))
nrow(fn125_tags)
head(fn125_tags)


filters <- list(lake = "HU", spc = "076", grp = "55")
fn125_tags <- get_FN125Tags(filters)
nrow(fn125_tags)
head(fn125_tags)
```


## FN125Lamprey - Observed Lamprey Wounds


FN125Lam records contain information about the individual lamprey
wounds observed on a sampled fish and can be fetched using the
`get_Fn125Lamprey()` function.  Historically, lamprey wounds were
reported as a single field (XLAM) in the FN125 table.  In the early
2000 the Great Lakes fishery community agreed to capture lamprey
wounding data in a more consistent fashion across the basin using the
conventions described in Ebener et al 2006.  The FN125Lam table
captures data from individual lamprey wounds collected using those
conventions.  A sampled fish with no observed wound will have a single
record in this table (with lamijc value of 0), while fish with lamprey
wounds, will have one record for every observed wound.  This function
takes an optional filter list which can be used to return records
based on several different attributes of the wound (wound type, degree
of healing, and wound size) as well as, attributes of the sampled fish
such as the species, or group code, or attributes of the effort, the
sample, or the project(s) that the samples were collected in.


```{r fn125_lamprey}
fn125_lam <- get_FN125Lam(list(
  lake = "ON",
  spc = "081",
  year = "2015",
  gear = "GL"
))
nrow(fn125_lam)
head(fn125_lam)


fn125_lam <- get_FN125Lam(list(
  lake = "SU", spc = "081", year__gte = 2015,
  lamijc_type = c("A1", "A2", "A3")
))
nrow(fn125_lam)
head(fn125_lam)



filters <- list(
  lake = "SU",
  prj_cd = c("LSA_IA15_CIN", "LSA_IA17_CIN"),
  eff = "051",
  spc = "091"
)
fn125_lam <- get_FN125Lam(filters)
nrow(fn125_lam)
head(fn125_lam)


filters <- list(lake = "HU", spc = "076", grp = "55")
fn125_lam <- get_FN125Lam(filters)
nrow(fn125_lam)
head(fn125_lam)
```


## Fn126 - Diet Data

The `get_Fn126()` function can be used to access the api endpoint to
for FN126 records. FN126 records contain the counts of identifiable
items in found in the stomachs of fish sampled and processed in the
field (the FN126 table does include more detailed analysis that is
often conducted in the laboratory).  The `get_FN126()` function takes
an optional filter list which can be used to return records based on
several different attributes of the diet item (taxon, taxon__like), as
well as, attributes of the sampled fish such as the species, or group
code, or attributes of the effort, the sample, or the project(s) that
the samples were collected in.

```{r fn126}
fn126 <- get_FN126(list(lake = "ON", year = 2012, spc = "334", gear = "GL"))
nrow(fn126)
head(fn126)



filters <- list(lake = "SU", prj_cd = c("LSA_IA12_CIN", "LSA_IA17_CIN"))
fn126 <- get_FN126(filters)
nrow(fn126)
head(fn126)

filters <- list(lake = "HU", spc = "076", grp = "55")
fn126 <- get_FN126(filters)
nrow(fn126)
head(fn126)
```


## Fn127 - Age Estimates

The `get_fn127()` function can be used to access the api endpoint to
for FN127 records which contain age estimate/interpretations.  This
function takes an optional filter list which can be used to return
records based on several different attributes of the age estimate such
as the assigned age, the aging structure, confidence, number of
complete annuli and edge code, or whether or not it was identified as
the 'preferred' age for this fish. Additionally, filters can be
applied to select age estimates based on attributes of the sampled
fish such as the species, or group code, or attributes of the effort,
the sample, or the project(s) that the samples were collected in.

```{r fn127}
fn127 <- get_FN127(list(lake = "ON", year = 2012, spc = "334", gear = "GL"))
nrow(fn127)
head(fn127)

filters <- list(
  lake = "ER",
  protocol = "TWL",
  spc = c("331", "334"),
  year = 2010,
  sidep__lte = 20
)
fn127 <- get_FN127(filters)
nrow(fn127)
head(fn127)

filters <- list(
  lake = "SU",
  prj_cd = c("LSA_IA15_CIN", "LSA_IA17_CIN"),
  eff = "051",
  spc = "091"
)
fn127 <- get_FN127(filters)
nrow(fn127)
head(fn127)



filters <- list(lake = "HU", spc = "076", grp = "55")
fn127 <- get_FN127(filters)
nrow(fn127)
head(fn127)
```


# Creel Data

The creel portal houses data from creel surveys that where collected
using the FN-II data model and exposes an api that make that data
available in a format that is very similar (if not identical to FN-II
tables).  The glfishr contains a number of function to fetch creel
data that are direct analogs to their fisheries assessment
counterparts.  Most of examples above will work by just changing the
function name from get_FN* to get_SC* and ensuring that any project
codes reference existing creel surveys.

As a simple example, here are several blocks of code that fetch the
data for a single creel and prints out the first few rows of each
table.  (Note that the filter is composed once, and used for all
subsequent functions).


##  SC011 - Creel Meta-data

```{r sc011}
creel_filter <- list(prj_cd = "LHA_SC08_033")
dat <- get_SC011(creel_filter)
dat <- anonymize(dat)
nrow(dat)
head(dat)
```

##  SC022 - Season Strata

```{r sc022}
dat <- get_SC022(creel_filter)
dat <- dat[with(dat, order(PRJ_CD, SSN)), ]
nrow(dat)
head(dat)
 ```


##  SC023 - Day Type Strata

```{r sc023}
dat <- get_SC023(creel_filter)
dat <- dat[with(dat, order(PRJ_CD, SSN, DTP)), ]
nrow(dat)
head(dat)
 ```

##  SC024 - Period Strata

```{r sc024}
dat <- get_SC024(creel_filter)
dat <- dat[with(dat, order(PRJ_CD, SSN, DTP, PRD)), ]
nrow(dat)
head(dat)
 ```

##  SC025 - Exception Dates

```{r sc025}
dat <- get_SC025(creel_filter)
nrow(dat)
head(dat)
 ```

##  SC026 - Spatial Strata

```{r sc026}
dat <- get_SC026(creel_filter)
dat <- dat[with(dat, order(PRJ_CD, SPACE)), ]
nrow(dat)
head(dat)
 ```

##  SC028 - Fishing Modes

```{r sc028}
dat <- get_SC028(creel_filter)
dat <- dat[with(dat, order(PRJ_CD, MODE)), ]
nrow(dat)
head(dat)
 ```


##  SC111 - Creel Logs

```{r sc111}
dat <- get_SC111(creel_filter)
nrow(dat)
head(dat)
 ```

##  SC112 - Activity Counts


```{r sc112}
dat <- get_SC112(creel_filter)
nrow(dat)
head(dat)
 ```

##  SC121 - Creel Interviews

```{r sc121}
dat <- get_SC121(creel_filter)
nrow(dat)
head(dat)
 ```

##  SC123 - Catch Counts

```{r sc123}
dat <- get_SC123(creel_filter)
nrow(dat)
head(dat)
 ```

##  SC125 - Biological Samples

```{r sc125}
dat <- get_SC125(creel_filter)
nrow(dat)
head(dat)
 ```


##  SC125 Lamprey - Observed Lamprey Wounds

```{r sc125Lam}
dat <- get_SC125Lam(creel_filter)
nrow(dat)
head(dat)
 ```

##  SC125 Tags - Applied or Recovered Tags

```{r sc125Tag}
# note the filter has changed
dat <- get_SC125Tags(list(year = 2000, lake = "HU"))
nrow(dat)
head(dat)
 ```
